---
title: "Kumi Michael Osei"
output: pdf_document
date: "2023-10-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(pander)
library(tidyverse)
library(knitr)
library(ggplot2)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
cbPalette <- c("#CC79A7", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# The palette with black:
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```


# Part A

Ten parts were selected randomly from a line and duplicate measurements of the part’s wall thickness were taken by each of three operators of the measurement apparatus. Click on the link and you will see the data in your browser.

```{r}
operator<-c(1 ,2 ,3)
op_rep<-rep(operator, each =2)
op_rep1<-seq(1,length(op_rep), by = 1)
op_rep2<-c(1,2)
var<-rep("op", length(op_rep))
lab<-c("obs", paste(var,op_rep,op_rep1, sep = "."))
dat1<-read.csv("partA_data.csv", header = F )
colnames(dat1)<-c(lab)
dat<-dat1 %>%
  gather("obs_code","value",op.1.1:op.3.6)%>%
  mutate(obs=as.numeric(obs))%>%
  mutate(value=as.numeric(value))
dat<-dat[order(dat$obs), ]
dat$operator<-rep(op_rep, times = length(dat1$obs))
dat$replicate<-rep(op_rep2, times = 3*length(dat1$obs))
dat$part<-rep(seq(1,10,by =1), each = length(op_rep1))
tidy_data<-(dat[, c(3,4,5,6)])
rownames(tidy_data) <- NULL
tidy_data<-tidy_data %>%
  mutate(operator=as.factor(operator))%>%
  mutate(replicate=as.factor(replicate))%>%
  mutate(part=as.factor(part))
head<-head(tidy_data,15)
kable(head,caption="Tidy data for the duplicate 
      measurements of part’s wall thickness by
      three operators for 10 parts")
```
```{r, message=FALSE}
#Averaging the replicates of each operator using the mean
meanrep<-rep(seq(1,length(tidy_data$operator)/2), each = 2)
tidy_data$meanrep<-as.factor(meanrep)
library(dplyr)

tidy2 <- tidy_data %>%
  group_by(meanrep, operator, part) %>%
  summarize(mean_value = mean(value))

head<-head(tidy2,15)
kable(head,caption="Summarized data with mean of replicates")
```

```{r}
library(ggplot2)
tidy2 %>% 
  ggplot(aes(x=part, y=tidy2$mean_value-(min(tidy2$mean_value))+0.001,
             fill =part)) +geom_col()+
  facet_wrap(~ operator)+
  labs(caption = "We have a histogram of the parts
       and their mean wall thickness. Each facet
       represents the operator used ", 
       y = "mean wall thickness(scale = y-min(y)+0.001)")
```

## Issues

The data had to scaled for the best visualization because it all the measurements were very close.

# Part B

Brain weight (g) and body weight (kg) for 62 species.

```{r}
mat<-as.matrix(read.csv("partB_data.csv", header = F ))
colnames(mat) <- NULL
vec<-numeric()
for(i in 1:nrow(mat)){
  vec <- c(vec, mat[i, ])
}
vec<-vec[complete.cases(vec)]
type<-c("Body_Wt","Brain_Wt")
M.type<-rep(type, times = length(vec)/2)
rep<-seq(1,length(vec)/2, by =1)
Species<-rep(rep,each =2)
data<-data.frame(vec,M.type,Species)
tidy_data<-spread(data, M.type, vec)
tidy_data<-tidy_data %>%
  mutate(Species=as.factor(Species))
head<-head(tidy_data,15)
kable(head,caption="Tidy data for Brain weight (g) 
      and body weight (kg) for 62 species")
```

```{r, include=FALSE}
zones <- matrix(c(1,1,1, 
                  0,5,0, 
                  2,6,4, 
                  0,3,0), ncol = 3, byrow = TRUE)
layout(zones, widths=c(0.3,4,1), heights = c(1,3,10,.75))
layout.show(n=6)
x=tidy_data$Body_Wt
y=tidy_data$Brain_Wt
scatterhist <- function(x, y, xlab = "", ylab = "", plottitle="", 
                        xsize=1, cleanup=TRUE,...){
  def.par <- par(no.readonly = TRUE)
  zones <- matrix(c(1,1,1, 0,5,0, 2,6,4, 0,3,0), ncol = 3, byrow = TRUE)
  layout(zones, widths=c(0.3,4,1), heights = c(1,3,10,.75))
  xhist <- hist(x, plot = FALSE)
  yhist <- hist(y, plot = FALSE)
  top <- max(c(xhist$counts, yhist$counts))
  par(xaxt="n", yaxt="n",bty="n",  mar = c(.3,2,.3,0) +.05)
  plot(x=1,y=1,type="n",ylim=c(-1,1), xlim=c(-1,1))
  text(0,0,paste(plottitle), cex=2)
  plot(x=1,y=1,type="n",ylim=c(-1,1), xlim=c(-1,1))
  text(0,0,paste(ylab), cex=1.5, srt=90)
  plot(x=1,y=1,type="n",ylim=c(-1,1), xlim=c(-1,1))
  text(0,0,paste(xlab), cex=1.5)
  par(mar = c(2,0,1,1))
  barplot(yhist$counts, axes = FALSE, xlim = c(0, top),
          space = 0, horiz = TRUE)
  par(mar = c(0,2,1,1))
  barplot(xhist$counts, axes = FALSE, ylim = c(0, top), space = 0)
  par(mar = c(2,2,.5,.5), xaxt="s", yaxt="s", bty="n")
  plot(x, y , pch=19, col="#00000022", cex=xsize, ...)
  if(cleanup) {par(def.par)}
}
```


```{r}
scatterhist(x, y, xlab="Body weight",
            ylab="Brain weight",   
  "plot for the brain weight and body weight fot the 62 species",
  xsize =2)
```

## Issues

The outliers in the data set interfered with the clarity of our scatter plot especially for our lower values.

# Part C

Gold Medal performance for Olympic Men’s Long Jump, year is coded as 1900=0. Goodness, ragged arrays. Check out fread in the data.table package.

```{r}
mat<-as.matrix(read.csv("partC_data.csv", header = F ))
colnames(mat) <- NULL
vec<-numeric()
for(i in 1:nrow(mat)){
  vec <- c(vec, mat[i, ])
}
vec<-vec[complete.cases(vec)]
dat<-data.frame(matrix(vec, ncol = 2, byrow = T))
dat$X1<-dat$X1+1900
colnames(dat)<-c("year", "Jump")
dat<-dat[order(dat$year), ]
tidy_data<-dat
rownames(tidy_data) <- NULL
tidy_data<-tidy_data %>%
  mutate(year=as.factor(year))
head<-head(tidy_data,15)
kable(head,caption="Tidy data  of  performance for 
      Olympic Men’s Long Jump and year")
```

```{r}
plot(x=tidy_data$year, y=tidy_data$Jump, 
     main = "Plot of Olympic Men’s Long Jump and year", 
     ylab="Long Jump", pch = 19, xlab ="Year")
```

## Issues

Data description was not very clear on the unit of measurement for the long jump.


# Part D

Triplicate measurements of tomato yield for two varieties of tomatoes at three planting densities.

```{r}
mat<-as.matrix(read.csv("partD_data.csv", header = F ))
colnames(mat) <- NULL
vec<-numeric()
for(i in 1:nrow(mat)){
  vec <- c(vec, mat[i, ])
}
yield<-vec
Variety<-c("Ife","PusaED")
Variety<-rep(Variety, each = length(vec)/2)
densities<-rep(c("10000","20000","30000"),each = 3)
densities<-rep(densities, times = 2)
tidy_data<-data.frame(yield,Variety,densities)
head<-head(tidy_data,15)
kable(head,caption="Tidy data for the triplicate measurements
      of tomato yield for two varieties of tomatoes 
      at three planting densities.")
```

```{r, message=FALSE}
# Mean of triplicate measurements
meanrep<-rep(seq(1,length(tidy_data$densities)/3), each = 3)
tidy_data$meanrep<-as.factor(meanrep)
tidy2 <- tidy_data %>%
  group_by(meanrep, densities, Variety) %>%
  summarize(mean_value = mean(yield))

head<-head(tidy2,15)
kable(head,caption="summarized tidy data with mean of
      triplicates of each density")
```


```{r}
tidy2 %>% 
  ggplot(aes(x=Variety, y=mean_value)) +
  geom_col(aes(fill=Variety)) + 
  facet_wrap(~ densities)+
  labs(caption = "We have a histogram of three
  planting densities and their yield.
       Each facet represents the variety")+
  scale_fill_manual(values=cbPalette)
```

## Issues 

The data did not have any issues.


# Part E

Larvae counts at two ages given 5 different treatments in 8 blocks.

```{r}
mat<-as.matrix(read.csv("partE_data.csv", header = F ))
colnames(mat) <- NULL
mat<-mat[,2:11]
vec<-numeric()
for(i in 1:nrow(mat)){
  vec <- c(vec, mat[i, ])
}
counts<-vec
b<-seq(1,8)
block<-rep(b, each = 10)
A<-rep(c(1,2), each = 5)
Age<-rep(A, times = 8)
t<-seq(1,5)
treatment<-rep(t, times = 8)
dat<-data.frame(counts, block, treatment, Age)
tidy_data<-dat %>%
  mutate(block=as.factor(block))%>%
  mutate(treatment=as.factor(treatment))
head<-head(tidy_data,15)
kable(head,caption="Tidy data for our larvae counts 
      at two ages given 5 different treatments in 8 blocks")
```

```{r}
# To use for fills, add
library(ggplot2)
tidy_data %>% 
  ggplot(aes(x=block, y=counts)) +
  geom_col(aes(fill=block)) + 
  facet_wrap(~ treatment)+
  scale_fill_manual(values=cbPalette)+
  labs(caption = "We have a histogram of blocks and 
       their larvae counts. Each facet represents 
       the Treatments. Treatment 1 to 5 for Age 1 
       and the rest for age 2")
```

## Issues

No Issues with data.
